\documentclass[12pt,letterpaper, onecolumn]{exam}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{array}
\usepackage[dvipsnames]{xcolor}
\usepackage[lmargin=71pt, tmargin=1.2in]{geometry}
\lhead{Mustafa Rashid\\}
\rhead{Chapters 5.2 \& 5.3\\}
\chead{\hline} 
\thispagestyle{empty} 
\newcommand*{\setdef}[1]{\left\{#1 \right\}} 
\newcommand{\doesnotdivide}{\not\hspace{2.5pt}\mid}

\begin{document}
	
	\begingroup  
	\noindent\LARGE Discrete Mathematics\\
	\noindent\LARGE Chapters 5.2 \& 5.3 Homework\\
	\noindent\large \today\\
	\noindent\large Mustafa Rashid\par
	\noindent\large Fall 2024\par
	\endgroup
	\rule{\textwidth}{0.4pt}
	\pointsdroppedatright
	\printanswers
	\renewcommand{\solutiontitle}{\noindent\textbf{Ans:}\enspace}  
	
	\centering\large Exercise Set 5.2\\
	\begin{questions}
		\setcounter{question}{1} \question  Use mathematical induction to show that any postage of at least 12$\mbox{\textcentoldstyle}$ can be obtained using 3$\mbox{\textcentoldstyle}$   and 7$\mbox{\textcentoldstyle}$  stamps.
		
		\setcounter{question}{3}\question  For each integer $n$ with $n\geq2$, let $P(n)$ be the formula
		$$\sum_{i=1}^{n-1}i(i+1)=\frac{n(n-1)(n+1)}{3}$$
		\begin{parts}
			\part Write $P(2)$. Is $P(2)$ true?
			\part Write $P(k)$.
			\part Write $P(k+1)$.
			\part In a proof by mathematical induction that the formula holds for all integers $n\geq2$, what must be shown in the inductive step?
		\end{parts}
		Prove by mathematical induction. Do not derive from Theorem 5.2.2 or Theorem 5.2.3.
		\setcounter{question}{6}\question For all integers $n\geq1$,
		$$1+6+11+16+...+(5n-4)=\frac{n(5n-3)}{2}.$$
		Prove by mathematical induction.
		\setcounter{question}{11}\question $$\frac{1}{1\cdot2}+\frac{1}{2\cdot3}+...+\frac{1}{n(n+1)}=\frac{n}{n+1}, \textrm{ for all integers $n\geq1$}$$
		Use the formula for the sum of the first $n$ integers and/or the formula for the sum of a geometric sequence to evaluate the sums in 23 \& 29.
		\setcounter{question}{22}\question $$7+8+9+10+...+600$$
		\setcounter{question}{28}\question $$1-2+2^2-2^3+...+(-1)^n2^n, \textrm{ where $n$ is a positive integer}$$
		Find the mistake in the proof fragment below.
		\setcounter{question}{34}\question \textbf{Theorem:} For any integer $n\geq1,$
		$$\sum_{i=1}^{n}i(i!)=(n+1)!-1.$$
		\textbf{"Proof (by mathematical induction): Let the property} $P(n) \textrm{ be } \sum_{i=1}^{n}i(i!)=(n+1)!-1$\\
		\textbf{Show that $P(1)$ is true:} When $n=1$
		$$\sum_{i=1}^{1}i(i!)=(1+1)!-1$$
		So $1(1!)=2!-1$\\
		and $1=1$\\
		Thus $P(1)$ is true."\\
			\end{questions}
		\centering\large Exercise Set 5.3\\
		\begin{questions}
		 \setcounter{question}{1}\question Experiment with computing values of the product $(1+\frac{1}{2})+(1+\frac{1}{3})...(1+\frac{1}{n})$ for small values of $n$ to conjecture a formula for this product for general $n$. Prove your conjecture by mathematical induction.
		 \setcounter{question}{4}\question Evaluate the sum $\sum_{k=1}^{n}\frac{k}{(k+1)!}$ for $n=1,2,3,4,$ and 5.\\
		 Make a conjecture about a formula for this sum for general $n$, and prove your conjecture by mathematical induction.
		 \setcounter{question}{11}\question For any integer $n\geq0, 7^n-2^n$ is divisible by 5.
		 \setcounter{question}{27}\question Prove that for all integers $n\geq1,$
		 $$\frac{1}{3}=\frac{1+3}{5+7}=\frac{1+3+5}{7+9+11}=...$$
		 $$=\frac{1+3+...+(2n-1)}{(2n+1)+...+(4n-1)}.$$
		 \setcounter{question}{34}\question Let $m$ and $n$ be any integers that are greater than or equal to 1.
		 \begin{parts}
		 	\part Prove that a necessary condition for an $m\times n$ checkerboard to be completely coverable by L-shaped trominoes is that $mn$ be divisible by 3.
		 	\part Prove that having $mn$ divisible by 3 is not a sufficient condition for an $m\times n$ checkerboard to be completely coverable by L-shaped trominoes.
		 \end{parts}
		\end{questions}


\end{document}