\documentclass[12pt,letterpaper, onecolumn]{exam}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{array}
\usepackage[dvipsnames]{xcolor}
\usepackage[lmargin=71pt, tmargin=1.2in]{geometry}
\lhead{Mustafa Rashid\\}
\rhead{Notes\\}
\chead{\hline} 
\thispagestyle{empty} 
\newcommand*{\setdef}[1]{\left\{#1 \right\}} 
\newcommand{\doesnotdivide}{\not\hspace{2.5pt}\mid}

\begin{document}

\begingroup  
\noindent\LARGE Discrete Mathematics\\
\noindent\large \today\\
\noindent\large Mustafa Rashid\par
\noindent\large Fall 2024\par
\endgroup
\rule{\textwidth}{0.4pt}
\pointsdroppedatright
\printanswers
\renewcommand{\solutiontitle}{\noindent\textbf{Ans:}\enspace}  


	\section{General Recursive Definitions and Structural Induction}
	Recursively defined sets 
	\begin{enumerate}
		\item BASE: A statement that certain objects belong to the set.
		\item RECURSION: A collection of rules indicating how to form new set objects from those already known to be in the set.
		\item RESTRICTION: A statement that no objects belong to the set other than those coming from 1 and 2.
	\end{enumerate}
	\textbf{A string over S}: Let S be a finite set with at least one element. A string over S is a finite sequnce of elemnts from S. The elements of S are \textbf{characters} of the string, and the \textbf{length} of a string is the number of characters it contains. The \textbf{null string over S} is defined to be the ``string" with no characters. It is usually denoted $\epsilon$ and is said to have length 0.\\
		\textbf{Structural Induction for Recursively Defined Sets}\\
		Let S be a set that has been defined recursively, and consider a property that objects in S may or may not satisfy. To prove that every object in S satisfies the property:
		\begin{enumerate}
			\item Show that each object in the BASE for S satisfies the property;
			\item Show that for each rule in the RECURSION, if the rule is applied to objects in S that satisfy the property, then the objects defined by the rule also satisfy the property.\\
			NOTE: Because no objects other than those obtained through the BASE and RECURSION conditions are contained in S, it must be the case that every object in S satisfies the property.
			\end{enumerate}
			\textbf{Recursive Function:} A function is said to be defined recursively or to be a recursive function if its rule of definition refers to itself. Because of this self-reference, it is sometimes difficult to tell whether a given recursive function is well defined.

\end{document}